---
output:
  word_document: default
  html_document: default
---
# BAN 502 - Clustering Assignment

## Khayrayyah Haamid-Day

```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(cluster) #algorithms for clustering
library(factoextra)
trucks <- read_csv("trucks.csv")
#summary(trucks)
```

```{r}
ggplot(trucks, aes(Distance, Speeding)) +
  geom_point(alpha = 0.4) + theme_bw()
```
*Task 1*

It does appear that there is natural clustering  happening. The first cluster has a stronger concentration of data around speeding of 12. The second cluster is more spread apart with data points extending from 0 to 100 speeding, though there is a strong concentration  below 25. Distance for cluster 1 is within a range of more than 0 and less than 75. While the distance for cluster 2 is mainly focused around 150 to 225.

```{r}
#Task 2

trucks_cleaned = recipe(~ Distance + Speeding, trucks)

trucks_dummy = trucks_cleaned %>% 
  step_scale(all_numeric()) %>%
  step_center(all_numeric())

trucks_dummy = prep(trucks_dummy, trucks)
trucks_cleaned = bake(trucks_dummy, trucks)

#summary(trucks_dummy)
```

```{r}
#Task 3
set.seed(64)
kclust = kmeans(trucks_cleaned, centers = 2)
#kclust
trucks_cleaned = augment(kclust, trucks_cleaned)
#trucks_cleaned

ggplot(trucks_cleaned, aes(Distance, Speeding, color = .cluster)) +
  geom_point(alpha = 0.4) + theme_bw()
```

The clusters look the same as the graph done before. 

```{r}
#Task 4
set.seed(412)
clusts = 
  tibble(k = 1:8) %>% 
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts

clusters =
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))

p1 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```

A value of 4 looks to be the most appropriate for this data.

```{r}
#Task 5 
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point() + theme_bw()
```

It appears 4 is the best.

```{r}
#Task 6
set.seed(64)
clusters = kmeans(trucks_cleaned, centers = 4)
#clusters
trucks = augment(clusters, trucks)

ggplot(trucks, aes(x=Distance,y=Speeding,color=factor(.cluster))) + geom_point() 
```
Using 4 clusters looks visually good and separated in a way that makes sense. 1 and 4 are clustered for lower speed while 2 and 3 are clustered with higher speeds. 2 and 4 are separated from 1 and 3 to account for larger distance differences. 
